Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change

### 2018-09-27 LatexNotes/Notes.tex
NM

\subsubsection{Final formulation}
Equation \eqref{varKdVugly} is essentially in the form we want to use, except that the notation is bad. In particular, we will want to reserve the symbol $H$ for the Hamiltonian. So making superficial changes to the notation gives
\begin{align}
2 \beta u_s + 3 \eps D^{-7/4} \, u u_x
+ \frac{4 \pi^2 \delta^2}{3 M^2} D^{1/2} \, u_{xxx} = 0
\end{align}
Importantly, $u(x,s)$ is $2\pi$-periodic in $x$ and has normalized energy $\mean{u^2} = 1$.
The definitions of each of the dimensionless variables in terms of the original, physical ones (denoted with a bar) are
\begin{align}
& u = D^{1/4} \, \bar{\eta}/a \\
& x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
& s = 2 \pi \beta M^{-1} \lambda_0^{-1} \bar{x}
\end{align}
A list of the fixed parameters is
\begin{align}
& \lambda_0 = \sqrt{g h_0}/f_p \quad \text{(upstream wavelength)} \\
& \eps = a / h_0 \quad \text{(dimensionless ratio of wave amplitude to depth)} \\
& \delta = h_0 / \lambda_0 \quad \text{(dimensionless ratio of depth to upstream wavelength)} \\
& \beta	\quad \text{(an unspecified small parameter, usually $\eps$ or $\delta^2$)}
\end{align}
Here is a list of the parameters that vary in space due to the fact that depth $d$ varies in space
\begin{align}
& c = \sqrt{gd} \quad \text{(the wavespeed as it varies in space)} \\
&\lambda_c = M \lambda_p = M \sqrt{gd}/f_p \quad \text{(characteristic wavelength as it varies in space)} \\
&D = d/h_0 \quad \text{(dimensionless depth ratio as it varies in space)} \\	
\end{align}
Note, in the experiments, $\delta = 0.33$ and $\eps$ lies in the range $2.4 \times 10^{-3}$--$2.4 \times 10^{-2}$.

\np
\subsection{Hamiltonian structure}
We will be using the truncated KdV equations (TKDV) with variable depth, in particular an abrupt depth change (ADC). The upstream depth is normalized to $1$ and the downstream (dimensionless) depth is $D_0$. Accordingly, the upstream and downstream Hamiltonians are given by
\begin{align}
\label{Hamiltonian}
& H^- = E_0^{1/2} H_3 - H_2 \\
& H^+ = D_0^{-13/4} E_0^{1/2} H_3 - D_0^{3/2} H_2 \\
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Numerical Methods
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}

>>>>

\subsubsection{Final formulation}
Equation \eqref{varKdVugly} is essentially in the form we want to use, except that the notation is bad. In particular, we will want to reserve the symbol $H$ for the Hamiltonian. So making superficial changes to the notation gives
\begin{align}
\label{varKdV1}
2 \beta u_s + 3 \eps D^{-7/4} \, u u_x
+ \frac{4 \pi^2 \delta^2}{3 M^2} D^{1/2} \, u_{xxx} = 0
\end{align}
Importantly, $u(x,s)$ is $2\pi$-periodic in $x$ and has normalized energy $\mean{u^2} = 1$.
The definitions of each of the dimensionless variables in terms of the original, physical ones (denoted with a bar) are
\begin{align}
& u = D^{1/4} \, \bar{\eta}/a \\
& x = 2\pi \lambda_c^{-1} (\bar{x} - c \bar{t})  \\
& s = 2 \pi \beta M^{-1} \lambda_0^{-1} \bar{x}
\end{align}
A list of the fixed parameters is
\begin{align}
& \lambda_0 = \sqrt{g h_0}/f_p \quad \text{(upstream wavelength)} \\
& \eps = a / h_0 \quad \text{(dimensionless ratio of wave amplitude to depth)} \\
& \delta = h_0 / \lambda_0 \quad \text{(dimensionless ratio of depth to upstream wavelength)} \\
& \beta	\quad \text{(an unspecified small parameter, usually $\eps$ or $\delta^2$)}
\end{align}
Here is a list of the parameters that vary in space due to the fact that depth $d$ varies in space
\begin{align}
& c = \sqrt{gd} \quad \text{(the wavespeed as it varies in space)} \\
&\lambda_c = M \lambda_p = M \sqrt{gd}/f_p \quad \text{(characteristic wavelength as it varies in space)} \\
&D = d/h_0 \quad \text{(dimensionless depth ratio as it varies in space)} \\	
\end{align}
Note, in the experiments, $\delta = 0.33$ and $\eps$ lies in the range $2.4 \times 10^{-3}$--$2.4 \times 10^{-2}$.

Let's make one more cosmetic change. Divide Eq.~\eqref{varKdV1} by $3 \delta^2$ and set $\beta = 3\delta^2/2$ to get
\begin{align}
\label{varKdV}
u_s + \eps \delta^{-2} D^{-7/4} \, u u_x
+ \mu D^{1/2} \, u_{xxx} = 0
\end{align}
where
\begin{equation}
\mu = \frac{4 \pi^2}{9 M^2}
\end{equation}

Then the Hamiltonian structure is given by
\begin{align}
\label{Hamiltonian}
& H^- = \eps \delta^{-2} H_3 - \mu H_2 \\
& H^+ = \eps \delta^{-2} D_0^{-7/4} H_3 - \mu D_0^{1/2} H_2
\end{align}

\np
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Numerical Methods
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}
